Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-11-10 text/ms.tex
Edit abstract

Black swans are statistically improbable events that nonetheless occur---often with profound consequences. While extremes in the physical environment, such as monsoons and heat waves, are widely studied and increasing in magnitude and frequency, it remains unclear the extent to which ecological populations buffer or suffer from such extremes. Here, we develop a probability model to estimate the degree of heavy-tailedness (presence of black swans) in ecological process noise. We apply our model to \NPops~time series from around the world across \NOrders~taxonomic orders and seven classes. We find strong evidence of black swans, but they are rare occurring in \overallMinPerc--\overallMaxPerc\% of populations: most frequently for birds (\AvesRangePerc\%) followed by mammals (\MammaliaRangePerc\%), insects (\InsectaRangePerc\%), and fishes (\OsteichthyesRangePerc\%). Black swans were not explained by any life-history covariates, but tended to be driven by external perturbations such as climate, severe winters, predator or parasite cycles, and the combined effects of multiple factors. Black swans were more frequently detected for populations with longer time series and lower levels of process noise; for shorter and noisier time series, our simulations suggested black-swan dynamics are often not identifiable as such. The presence of black swans in population dynamics highlights the importance of developing robust conservation and management strategies---particularly as the frequency and magnitude of climatic extremes increase over the next century.

>>>>

Black swans are statistically improbable events that nonetheless occur---often with profound consequences. While extremes in the physical environment, such as monsoons and heat waves, are widely studied and increasing in magnitude and frequency, it remains unclear the extent to which ecological populations buffer or suffer from such extremes. Here, we develop a probability model to estimate the degree of heavy-tailedness (presence of black swans) in ecological process noise. We apply our model to \NPops~time series from around the world across \NOrders~taxonomic orders and seven classes. We find strong evidence of black swans, but they are rare occurring in \overallMinPerc--\overallMaxPerc\% of populations: most frequently for birds (\AvesRangePerc\%) followed by mammals (\MammaliaRangePerc\%), and insects (\InsectaRangePerc\%). Black swans were predominantly downward events (\percBSDown \%) and were not explained by any life-history covariates, but tended to be driven by external perturbations such as climate, severe winters, predator or parasite cycles, and the combined effects of multiple factors. Extreme events were more frequently detected for populations with longer time series and lower levels of process noise; for shorter and noisier time series, our simulations suggested black-swan dynamics are often not identifiable as such. The presence of black swans in population dynamics highlights the importance of developing robust conservation and management strategies---particularly as the frequency and magnitude of climate extremes increase over the next century.
